---
title: "p8105_hw1_ske2118_final"
author: "Shayne Estill"
date: "2024-09-21"
output: github_document
---

# Homework 1 Set up 

Here's a **code chunk** that loads tidyverse: 

```{r setup, include = FALSE}
library(tidyverse)
```


# Problem 1

Here we load our penguin data: 

```{r summary, include = FALSE}
data("penguins", package = "palmerpenguins")
summary(penguins)
```

The penguin dataset has observations about 
  three different penguin species (adelie, chinstrap, and gentoo) on 
  three different islands (biscoe, dream, and torgersen). 
  
Data collected about these penguins included 
  bill_length_mm, bill_depth_mm, and flipper_length_mm (all measured in millimeters)
  as well as body_mass_g (measured in grams), sex, and year. 
  
The penguins dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.



```{r scatterplot, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("scatterplot.pdf")
```
  
  I produced a scatterplot of bill_length_mm to flipper_length_mm in which
  each species is a different color point. I suppressed the warning           regarding two observations without a number value. 
  
# Problem 2 
```{r dataframe}
problem2_df = tibble(   
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric>0,
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H","I", "J"),
  vec_factor = factor(c("small", "medium", "large", "small", "medium",                    "large", "small", "medium", "large", "small"))
  )
  
```

Now, let's look at what happens when we try to take the means. 
```{r dataframe means}
  mean(pull(problem2_df, vec_numeric))
  mean(pull(problem2_df, vec_logical))
  mean(pull(problem2_df, vec_char))
  mean(pull(problem2_df, vec_factor))
```
  The vec_numeric and vec_logical are both able to produce mean values.
  The vec_char and vec_factor are NOT able to produce mean values. 

Now, let's try to explicitly convert variables from one type to another. 
```{r dataframe coercion}
  problem2_df|>
  mutate(
    vec_logical = as.numeric(vec_logical),
    vec_char = as.numeric(vec_char),
    vec_factor = as.numeric(vec_factor)
  )
```
When we applied the as.numeric function to our dataframe, we see that all vectors 
except for the character vector are able to be converted to numeric. 
This is because the factor variable was able to be successfully converted as for factor variables they can recognize the characters by their alphabetical order
and rank them in that way. However the character vector do not have any kind of factor
they can be ranked by and thus we see the NA output for vec_char. 

```{r dataframe coercion mean}
   mean(pull(problem2_df, vec_logical))
  mean(pull(problem2_df, vec_char))
  mean(pull(problem2_df, vec_factor))
```

This somewhat helps explain what happens when we try to take the mean in that 
we still are not able to calculate means for character and factor variables and
we receive NA errors. 