---
title: "p8105_hw1_ske2118_final"
author: "Shayne Estill"
date: "2024-09-21"
output: github_document

---


I'm an R Markdown document! 

# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.

#learning assessment#

```{r learning_assessment_1}
library(tidyverse)

la_df = tibble(
  norm_samp = rnorm(500, mean = 1),
  norm_samp_pos = norm_samp > 0,
  abs_norm_samp = abs(norm_samp)
)

ggplot(la_df, aes(x = abs_norm_samp)) + geom_histogram()

median_samp = median(pull(la_df, norm_samp))
```

The median of the variable containing absolute values is 
`r round(median_samp, digits = 2)`.

```{r}

```


#homework 1
{r setup, include = FALSE}
library(tidyverse)


#Problem 1
{r summary, include = FALSE}
data("penguins", package = "palmerpenguins")
summary(penguins)


The penguin dataset has observations about 
  three different penguin species (adelie, chinstrap, and gentoo) on 
  three different islands (biscoe, dream, and torgersen). 
  
Data collected about these penguins included 
  bill_length_mm, bill_depth_mm, and flipper_length_mm (all measured in millimeters)
  as well as body_mass_g (measured in grams), sex, and year. 
  
The penguins dataset has r nrow(penguins) rows and r ncol(penguins) columns.

The mean flipper length is r mean(penguins$flipper_length_mm, na.rm = TRUE).


##


{r scatterplot}
Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable (adding color = ... inside of aes in your ggplot code should help).

Export your first scatterplot to your project directory using ggsave.

ggplot(plot_penguin, aes(x = bill_length_mm, y = flipper_length_mm)) + geom_point()

  
